### 2021-11-24 draft.tex
Deploying to gh-pages from @ gottacatchenall/thesis_proposal@2056db570ad5def0d6b14696f23b9cb0d263a24f 🚀

%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}

@@

Within the last several hundred years, human activity has induced rapid
changes in Earth's atmosphere, oceans, and surface. Greenhouse gas
emissions have caused an increase the temperature of both Earth's
terrain and oceans, and both agricultural and urban development has
rapidly reshaped the Earth's land cover. These the bulk of this change
has occurred within the last several hundred years, a geological
instant, inducing a sudden shift in conditions to Earth's climate and
biosphere. As a result \emph{ecological forecasting}---modeling how
ecosystems and their services will change in the future---and then using
these forecasts to make decisions to mitigate the negative consequences
of this change on ecosystems, their functioning, and the services they
provide to humans has emerged as an imperative for ecology and
environmental science (Dietze 2017). However, robust prediction of
ecological processes is, to say the least, quite difficult (Beckage
\emph{et al.} 2011; Petchey \emph{et al.} 2015). This difficultly is
compounded by a few factors, the first being that sampling ecosystems is
not easy. Ecological data is often biased, noisey, and sparse in both
space and time. The current paucity of ecological data has resulted in
much interest in developing global systems for \emph{ecosystem
monitoring} (Makiola \emph{et al.} 2020), which would systematize the
collection of biodiversity data in manner that makes detecting and
predicting change more possible than at the moment (Urban \emph{et al.}
2021).

The second major challenge in ecological forecasting is that the
underlying dynamics of most ecological processes are unknown and instead
must be inferred from this (sparse) data. Much of the history of
quantitatively modeling ecosystems have been done in the language of
dynamical systems, describing how the value of an observable state of
the system, represented by a vector of numbers
\([x_1, x_2, \dots, x_n]^T = \vec{x}\) changes as over time, yielding
models in the form of differential equations in continuous-time
settings, \(\frac{dx}{dt} = f(x)\), or difference equations in
discrete-time settings, \(x_t = f(x_{t-1})\), where
\(f:\mathbb{R}^n \to \mathbb{R}^n\) is an arbitrary function describing
how the system changes on a moment-to-moment basis (e.g.~in the context
of communities, \(f\) could be Lotka-Voltera, Holling-Type-III or
DeAngelis-Beddington functional response). The form of this functional
response in real systems, and whether it is meaningfully non-zero for a
given species interaction, is effectively unknown and must be predicted
(Strydom \emph{et al.} 2021b), and some forms are inherently more
``forecastable'' than others (Beckage \emph{et al.} 2011; Chen \emph{et
al.} 2019; Pennekamp \emph{et al.} 2019). The initial success of these
forms of models can be traced back to the larger program of ontological
reductionism, which became the default approach to modeling in the
sciences after its early success in physics, which, by the time ecology
was becoming a quantitative science (sometime in the 20th century,
depending on who you ask), became the foundation for mathematical models
in ecology.

However, we run into many problems when aiming to apply this type of
model to empirical ecological data. Ecosystems are perhaps the
quintessential example of system that cannot be understood by iterative
reduction of its components into constituent parts---ecological
phenomena are emergent: the product of different mechanisms operating a
different spatial, temporal, and organizational scales (Levin 1992).
Further this analytical approach to modeling explicitly ignores known
realities: ecological dynamics not deterministic and many analytic
models in ecology assume long-run equilibrium. Finally, perhaps the
biggest challenge in using these models to describe ecological processes
is ecosystems consist of more dimensions than the tools of analytic
models are suited for. As the number of variables in an analytic model
increases, so does the ability of the scientist to discern clear
relationships between them given a fixed amount of data, the so-called
``curse'' of dimensionality.

But these problems are not solely unique to ecology. The term
\emph{ecological forecasting} implicitly creates an analogy with weather
forecasting. Although it has become a trite joke to complain about the
weather forecast being wrong, over the least 50 years the field of
numerical weather prediction (NWP) has dramatically improved out ability
to predict weather across the board (Bauer \emph{et al.} 2015). The
success of NWP, and the Earth observations systems that support it (Hill
\emph{et al.} 2004), should serve as a template for development of a
system for monitoring Earth's biodiversity. Much like ecology, NWP is
faced with high-dimensional systems that are governed by different
mechanisms at different scales. The success of NWP is that, rather than,
say, attempt to forecast the weather in Quebec by applying Navier-Stokes
to entire province, to instead use simulation models which describe
known mechanisms at different scales, and use the availability to
increasing computational power to directly simulate many batches of
dynamics which directly incorporate stochasticity and uncertainty in
parameter estimates via random number generation.

But forecasting is only half the story---if indeed ``{[}ecologists{]}
have hitherto only interpreted the world in various ways; the point is
to change it,'' then once we have a forecast about how an ecosystem will
change in the future, what if this forecast predicts a critical
ecosystem service will deteriorate? We are still left with the question,
what do we in the time being to mitigate the potentially negative
consequences a forecast predicts? In this framing, mitigating the
consequences of anthropogenic change on ecosystems becomes an
optimization problem: given a forecast of the future state of the
system, and some ``goal'' state for the future, the problem is then to
optimize our intervention into the system to maximize the probability
the system approaches our ``goal'' state. This dissertation aims to this
framework for ecosystem monitoring and forecasting
(fig.~\ref{fig:thesis}, left), and each chapter address some aspect of
this pipeline to data from a monitoring network to forecasts to
mitigation strategy (fig.~\ref{fig:thesis}, right).

@@

The primary research challenges this thesis addresses: how do we design
ecological samples to? How do we build the software infrastructure to
assimilate data from a variety of sources? How do we propagate
uncertainty from data to forecasts? The flow of chapters follows the
flow in fig.~\ref{fig:thesis} (left), from data collection via a
monitoring network, to forecasting an essential biodiversity variable
(EBV), to optimizing mitigation strategy based on constraints. In
chapter one, we discuss how simulation can aid in the design of
ecological samples and monitoring network design. In chapter two we use
data to forecast the uncoupling of a plant-pollinator network. In
chapter three, we apply simulation methods in landscape ecology to
optimize corridor placement with respect maximize the
time-until-extinction of a metapopulation. The fourth and final chapter
is the software (\emph{MetacommunityDynamics.jl}) which enables the rest
of the dissertation.

@@

This chapter uses simulation models to investigate the relationship
between species relative abundance, sampling effort, and probability of
observing an interaction between species in order to aid in the design
of samples of ecological interactions, and to provide a null expectation
of false-negative probability for a dataset of a given size. Further it
then proposes a method for optimizing the spatial sampling locations to
maximize the probability of detecting an interaction between two species
given a fixed number of total of observations and the distributions of
each species. This addresses the optimization of monitoring network part
of the flow from data to mitigation at the top of fig.~\ref{fig:thesis},
left. There is more than one way to observe a false-negative when
sampling interactions. fig.~\ref{fig:fnrtaxonomy} shows a taxonomy of
false-negatives in occurrence, co-occurrence, and interaction data. This
chapter untangles the relationship between co-occurrence false-negatives
and interaction false-negatives.

@@

We begin by proposing a method to compute a null expectation of the
probability of an interaction false-negative as a function number of
total observations of individuals of \emph{all species in the species
pool}. This is done by simulating the process of observation, where the
probability of observing a given species is its relative abundance. We
show that the realized false-negative rate can be quite high, simply as
a byproduct of the distributoin of relative abundance in communities. We
use a log-normal distribution of relative abundance (Hubbell 2001) and
simulating the process of observation on food-webs generated using the
niche model (Williams \& Martinez 2000) with connectance parameterized
by the flexible-links model (MacDonald \emph{et al.} 2020). An example
of this relation for networks with varying spceies richness is shown in
fig.~\ref{fig:fnr}.

@@

We then go on to testing some assumptions of this neutral model with
empirical data. Primarily that we analytically show that our neutral
model, if anything, underestimates the probability of false-negatives if
there are positive associations between species co-occurrence, and we
show these positive associations exist in two sets of spatially
replicated samples of interaction networks (Thompson \& Townsend 2000;
Hadfield \emph{et al.} 2014), fig.~\ref{fig:posassoc}---further I'm
planning to add the field data from the previous chapter into this
analysis once available. Finally this chapter proposes a simulated
annealing method to optimize the a set of \(n\) points in space to
maximize the probability of detecting an interaction between two species
\(a\) and \(b\) with \emph{known} distributions \(D_a\), \(D_b\).

@@

The first major result is using the simulation of the observation
process described above to generate expectations of interaction
false-negative rate (FNR) as a function of total number of observations,
with the goal being for this estimate to be used as correction for
detection error when fitting an interaction prediction model. This
relationship varies with the total richness of the metaweb
fig.~\ref{fig:fnr}.

@@

The data for this chapter is derived from multiple sources that can be
split into four categories. (1) Field data from three different field
sites across Colorado, each with multiple plots across an elevational
gradient, for seven, seven, and three years respectively. This data was
collected by Paul CaraDonna and Jane Oglevie (from the Rocky Mountain
Biological Laboratory; RMBL) and Julian Resasco (CU Boulder). (2) GBIF
spatial occurrence records of each of these species across Colorado,
including a metaweb of interactions across all of Colorado taken from
GBIF. (3) Remotely sensed data consisting of current and forecasting
bioclimatic variables from CHELSA. (4) Phylogenies for both bee and
flower species derived from NCBI GenBank barcodes for mitochondrial COI
(bumblebees) and chloroplast rbcL (flowers).

As the data we have is spatially sparse and likely to contain many
interaction ``false-negatives'' (Strydom \emph{et al.} 2021b), we begin
by predicting a metaweb of interactions across Colorado as they exist
\emph{in the present}. We do this using a set of candidate interaction
prediction models: relative abundance only, phylogenetic embedding only
(a la Strydom \emph{et al.} (2021a)), niche embedding only (Gravel
\emph{et al.} 2019), and all pairwise combinations of those constituent
models. After validating and selecting the best performing model, we
then predict how these distributions of each of these species will
change under the CMIP6 consensus climate forecast (Karger \emph{et al.}
2017), and then finally quantify the reduction in spatial between
species for which there is a predicted interaction.

@@

As land-use change has caused many habitats to become fragmented and
patchy, promoting landscape connectivity has become of significant
interest to mitigate the effects of this change on Earth's biodiversity.
However, the practical realities of conservation mean that there is a
limitation on how much we can modify landscapes in order to do this. So
what is the best place to put a corridor given a constraint on how much
surface-area you can change in a landscape? This is the question this
chapter seeks to answer. Models for inferring corridor locations have
been developed, but are limited in that are not developed around
promoting some element of ecosystem function, but instead by trying to
find the path of least resistance in an existing landscape from a
derived resistance surface (Peterman 2018). This chapter proposes a
general algorithm for choosing corridor placement to optimize a
measurement of ecosystem functioning derived from simulations run on
each proposed landscape modification.

@@

We propose various landscape modifications which alter the cover of a
landscape, represented as a raster. We then compute a new resistance
surface based on the proposed landscape modification using Circuitscape
(McRae \emph{et al.} 2008), and based on the values of resistance to
dispersal between pair of locations we simulate spatially-explicit
metapopulation dynamics model (Hanski \& Ovaskainen 2000; Ovaskainen
\emph{et al.} 2002) to estimate a distribution of time until extinction
for each landscape modification. The largest challenge in implementing
this algorithm is the space of potential modifications grows as
\(O((nm)!)\) for an \(n\) by \(m\) raster. For most actual landscapes to
which we wish to apply this method, the set of possible modifications
becomes uncomputably large, so we use simulated annealing to explore the
search space of possible modifications to estimate the modification that
maximizes the time-until extinction of simulated metapopulation dynamics
under that hypothetical modified landscape.

The biggest challenge in implementing simulated annealing in this
context is defining a proposal function for landscape modifications. At
the moment this is done by computing the minimum-spanning-tree (MST) of
the spatial nodes, and then proposing corridors that connect nodes that
are already connected in the MST. The primary reason for doing this is
the cut down the size of the solution space to enable quicker
convergence, although the final software that implements this algorithm
will enable alternative methods of proposing modifications.

The goal output of this chapter is not only to provide a set of discrete
corridor options, but also to rank the cells in the raster by priority
based on how many times they are converted in the distribution of
``good'' corridors after simulated annealing has reached a
pseudoequilibrium. Further, the final component of this chapter is
measuring the effect of land-use change on the robustness of the
optimized corridor by simulating various neutral models of urban and
agricultural sprawl, and determining if the proposed modifications still
maximize time-until-extinction when the landcover in the landscape is
not static.

@@

The current progress is that I have an algorithm for proposing landscape
modifications and a simple implementation of simulated annealing. The
only gap left is implementing Circuitscape estimation of resistance
surfaces.

@@

The final chapter consists of a collection of modules in the Julia
language for different aspects of community ecology, including most of
the code used for the preceding chapters. Indeed
\texttt{MetacommunityDynamics.jl} (MCD.jl) is the epicenter of this set
of tools, but due to the nature of the Julia language, MCD.jl is
interoperable with serveral existing packages within the
\texttt{EcoJulia} organization, including several to which I have
contributed. We need a software library like this to generate synthetic
data from a \emph{known} set of mechanisms and parameters to test our
methods for parameter inference and forecasting on this \emph{known}
system to assess the effectiveness of these inference and forecasting
methods.

@@

Below (fig.~\ref{fig:foodwebtraj}) is a sample output of simulated
food-web dynamics for a metaweb of 100 species generated using the
minimum-potential-niche model with connectance \(C=0.05\) and
forbidden-link probability of \(0.5\). The dynamics change according to
a Lotka-Volttera functional response, dispersal (with dispersal distance
inverse proportional to trophic-level), linear mortality, and logistic
growth for any species at the producer trophic-level.

@@

The software as it exists is capable of simulating the biomass dynamics
of arbitrarily large food-webs using Lotka-Volterra, Holling Type-II, or
Holling Type-III functional responses. It currently has methods to
implement Gaussian drift, and verious forms of dispersal via
Dispersal.jl. Also occupancy dynamics for Levins metapopulations (Levins
1969), and spatially explicit Hanski-Ovaskainen metapopulations (Hanski
\& Ovaskainen 2000; Ovaskainen \emph{et al.} 2002). This is most of what
needs to exist for the preceding chapters. In-progress functionality
includes selection (which affects growth-rate) on arbitrary
environmental variables in progress, as well as traits.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
Banville, F., Vissault, S. \& Poisot, T. (2021). Mangal.jl and
EcologicalNetworks.jl: Two complementary packages for analyzing
ecological networks in Julia. \emph{Journal of Open Source Software}, 6,
2721.

>>>>

\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%

@@

Within the last several hundred years, human activity has induced rapid
changes in Earth's atmosphere, oceans, and surface. Greenhouse gas
emissions have caused an increase the temperature of both Earth's
terrain and oceans, and both agricultural and urban development has
rapidly reshaped Earth's land cover. The bulk of this change has
occurred within the last several hundred years, a geological instant,
inducing a sudden shift in conditions to Earth's climate and biosphere.
As a result \emph{ecological forecasting}---modeling how ecosystems and
their services will change in the future---and then using these
forecasts to make decisions to mitigate the negative consequences of
this change on ecosystems, their functioning, and the services they
provide to humans has emerged as an imperative for ecology and
environmental science (Dietze 2017). However, robust prediction of
ecological processes is, to say the least, quite difficult (Beckage
\emph{et al.} 2011; Petchey \emph{et al.} 2015). This difficultly is
compounded by a few factors, the first being that sampling ecosystems is
not easy. Ecological data is often biased, noisey, and sparse in both
space and time. The current paucity of ecological data has resulted in
much interest in developing global systems for \emph{ecosystem
monitoring} (Makiola \emph{et al.} 2020), which would systematize the
collection of biodiversity data in a manner that makes detecting and
predicting change more possible than at the moment (Urban \emph{et al.}
2021).

The second major challenge in ecological forecasting is that the
underlying dynamics of ecological processes are unknown and instead must
be inferred from this (sparse) data. Much of the history of
quantitatively modeling ecosystems have been done in the language of
dynamical systems, describing how the value of an observable state of
the system, represented by a vector of numbers
\([x_1, x_2, \dots, x_n]^T = \vec{x}\) changes over time, yielding
models in the form of differential equations in continuous-time
settings, \(\frac{dx}{dt} = f(x)\), or difference equations in
discrete-time settings, \(x_t = f(x_{t-1})\), where
\(f:\mathbb{R}^n \to \mathbb{R}^n\) is an arbitrary function describing
how the system changes on a moment-to-moment basis (e.g.~in the context
of communities, \(f\) could be Lotka-Voltera, Holling-Type-II or
DeAngelis-Beddington functional response). The form of this functional
response in real systems, and whether it is meaningfully non-zero for a
given species interaction, is effectively unknown and must be predicted
(Strydom \emph{et al.} 2021b), and some forms of these dynamics are
inherently more ``forecastable'' than others (Beckage \emph{et al.}
2011; Chen \emph{et al.} 2019; Pennekamp \emph{et al.} 2019). The
initial success of these forms of models can be traced back to the
larger program of ontological reductionism, which became the default
approach to modeling in the sciences after its early success in physics,
which, by the time ecology was becoming a quantitative science (sometime
in the 20th century, depending on who you ask), became the foundation
for mathematical models in ecology.

However, we run into many problems when aiming to apply this type of
model to empirical ecological data. Ecosystems are perhaps the
quintessential example of system that cannot be understood by iterative
reduction of its components into constituent parts---ecological
phenomena are emergent: the product of different mechanisms operating at
different spatial, temporal, and organizational scales (Levin 1992).
Further this analytical approach to modeling explicitly ignores known
realities: ecological dynamics not deterministic and many analytic
models in ecology assume long-run equilibrium. Finally, perhaps the
biggest challenge in using these models to describe ecological processes
is ecosystems consist of more dimensions than the tools of analytic
models are suited for. As the number of variables in a model increases,
so does the ability of the scientist to discern clear relationships
between them given a fixed amount of data, the so-called ``curse'' of
dimensionality.

But these problems are not solely unique to ecology. The term
\emph{ecological forecasting} implicitly creates an analogy with weather
forecasting. Although it has become a trite joke to complain about the
weather forecast being wrong, over the least 50 years the field of
numerical weather prediction (NWP) has dramatically improved our ability
to predict weather across the board (Bauer \emph{et al.} 2015). The
success of NWP, and the Earth observation systems that support it (Hill
\emph{et al.} 2004), should serve as a template for development of a
system for monitoring Earth's biodiversity. Much like ecology, NWP is
faced with high-dimensional systems that are governed by different
mechanisms at different scales. The success of NWP is that, rather than,
say, attempt to forecast the weather in Québec by applying Navier-Stokes
to entire province, to instead use simulation models which describe
known mechanisms at different scales, and use the availability to
increasing computational power to directly simulate many batches of
dynamics which directly incorporate stochasticity and uncertainty in
parameter estimates via random number generation.

But forecasting is only half the story---if indeed ``{[}ecologists{]}
have hitherto only interpreted the world in various ways; the point is
to change it,'' then once we have a forecast about how an ecosystem will
change in the future, what if this forecast predicts a critical
ecosystem service will deteriorate? We are still left with the question,
what do we in the time being to mitigate the potentially negative
consequences a forecast predicts? In this framing, mitigating the
consequences of anthropogenic change on ecosystems becomes an
optimization problem: given a forecast of the future state of the
system, and some ``goal'' state for the future, the problem is then to
optimize our intervention into the system to maximize the probability
the system approaches our ``goal'' state. This dissertation aims to
provide a framework for ecosystem monitoring and forecasting
(fig.~\ref{fig:thesis}, left), and each chapter address some aspect of
this pipeline to data from a monitoring network to forecasts to
mitigation strategy (fig.~\ref{fig:thesis}, right).

@@

The primary research challenges this thesis addresses are: how do we
design ecological samples to? How do we build the software
infrastructure to assimilate data from a variety of sources? And how do
we propagate uncertainty from data to forecasts? The flow of chapters
follows the flow in fig.~\ref{fig:thesis} (left), from data collection
via a monitoring network, to forecasting an essential biodiversity
variable (EBV), to optimizing mitigation strategy based on constraints.
In chapter one, we discuss how simulation can aid in the design of
ecological samples and monitoring network design. In chapter two we use
data to forecast the uncoupling of a plant-pollinator network. In
chapter three, we apply simulation methods in landscape ecology to
optimize corridor placement to maximize the time-until-extinction of a
metapopulation. The fourth and final chapter is the software
(\emph{MetacommunityDynamics.jl}) which enables the rest of the
dissertation.

@@

This chapter uses simulation models to investigate the relationship
between species relative abundance, sampling effort, and probability of
observing an interaction between species in order to aid in the design
of samples of ecological interactions across space, and to provide a
null expectation of the false-negative probability for a dataset of a
given size. Further it then proposes a method for optimizing the spatial
sampling locations to maximize the probability of detecting an
interaction between two species given a fixed number of total of
observations and the distributions of each species. This addresses the
optimization of monitoring network part of the flow from data to
mitigation at the top of fig.~\ref{fig:thesis}, left. There is more than
one way to observe a false-negative when sampling interactions---
fig.~\ref{fig:fnrtaxonomy} shows a taxonomy of false-negatives in
occurrence, co-occurrence, and interaction data. The goal of this
chapter is to untangle the relationship between co-occurrence
false-negatives (and their relation to relative abundance) and
interaction false-negatives.

@@

We begin by proposing a method to compute a null expectation of the
probability of an interaction false-negative as a function number of
total observations of individuals of \emph{all species in the species
pool}. This is done by simulating the process of observation, where the
probability that an observation is of a given species is that species'
relative abundance. We show that the realized false-negative rate can be
quite high, simply as a byproduct of the distribution of relative
abundance in communities. We use a log-normal distribution of relative
abundance (Hubbell 2001) and simulating the process of observation on
food-webs generated using the niche model (Williams \& Martinez 2000)
with connectance parameterized by the flexible-links model (MacDonald
\emph{et al.} 2020). An example of this relation for networks with
varying species richness is shown in fig.~\ref{fig:fnr}.

@@

We then go on to test some assumptions of this neutral model with
empirical data. Primarily we analytically show that our neutral model,
if anything, underestimates the probability of false-negatives if there
are positive associations between species co-occurrence, and we show
these positive associations exist in two sets of spatially replicated
samples of interaction networks (Thompson \& Townsend 2000; Hadfield
\emph{et al.} 2014), fig.~\ref{fig:posassoc}---further I'm planning to
add the field data from chapter two into this analysis once available.
Finally this chapter proposes a simulated annealing method to optimize
the a set of \(n\) points in space to maximize the probability of
detecting an interaction between two species \(a\) and \(b\) with
\emph{known} distributions \(D_a\), \(D_b\).

@@

The first major result is using the simulation of the observation
process described above to generate expectations of interaction
false-negative rate (FNR) as a function of total number of observations,
with the goal being for this estimate to be used as correction for
detection error when fitting an interaction prediction model. This
relationship varies with the total richness of the metaweb (see
fig.~\ref{fig:fnr}).

@@

The data for this chapter is derived from multiple sources that can be
split into four categories. (1) Field data from three different field
sites across Colorado, each with multiple plots across an elevational
gradient, for seven, seven, and three years respectively. This data was
collected by Paul CaraDonna and Jane Oglevie (from the Rocky Mountain
Biological Laboratory; RMBL) and Julian Resasco (CU Boulder). (2) GBIF
spatial occurrence records of each of these species across Colorado,
including a metaweb of interactions across all of Colorado taken from
GBIF. (3) Remotely sensed data consisting of current and forecasted
bioclimatic variables from CHELSA. (4) Phylogenies for both bee and
flower species derived from NCBI GenBank barcodes for mitochondrial COI
(bumblebees) and chloroplast rbcL (flowers).

As the data we have is spatially sparse and likely to contain many
interaction false-negatives, we begin by predicting a metaweb of
interactions across Colorado as they exist \emph{in the present}. We do
this using a set of candidate interaction prediction models: relative
abundance only, phylogenetic embedding only (a la Strydom \emph{et al.}
(2021a)), niche embedding only (encompassing both phenology and
environmental niche, similar to Gravel \emph{et al.} (2019)), and all
pairwise combinations of those constituent models. After validating and
selecting the best performing model, we then predict how these
distributions of each of these species will change under the CMIP6
consensus climate forecast (Karger \emph{et al.} 2017), and then finally
quantify the reduction in spatial overlap between species for which
there is a predicted interaction.

@@

As land-use change has caused many habitats to become fragmented and
patchy, promoting landscape connectivity has become of significant
interest to mitigate the effects of this change on Earth's biodiversity.
However, the practical realities of conservation mean that there is a
limitation on how much we can modify landscapes in order to do this. So
what is the best place to put a corridor given a constraint on how much
surface-area you can change in a landscape? This is the question this
chapter seeks to answer. Models for inferring corridor locations have
been developed, but are limited in that are not developed around
promoting some element of ecosystem function, but instead by trying to
find the path of least resistance in an existing landscape from a
derived resistance surface (e.g. Peterman 2018). This chapter proposes a
general algorithm for optimizing spatial restoration effort to move a
measurement of ecosystem functioning toward a target state.

@@

We propose various landscape modifications that alter the cover of a
landscape, represented as a raster. We then compute a new resistance
surface based on the proposed landscape modification using Circuitscape
(McRae \emph{et al.} 2008), and based on the values of resistance to
dispersal between pair of locations we simulate spatially-explicit
metapopulation dynamics (Hanski \& Ovaskainen 2000; Ovaskainen \emph{et
al.} 2002) to estimate a distribution of time-until-extinction for each
landscape modification. The largest challenge in implementing this
algorithm is the space of potential modifications grows as \(O((nm)!)\)
for an \(n\) by \(m\) raster. For most actual landscapes to which we
wish to apply this method, the set of possible modifications becomes
uncomputably large, so we use simulated annealing to explore the search
space of possible modifications to estimate the modification that
maximizes the time-until extinction of simulated metapopulation dynamics
under that hypothetical modified landscape.

The biggest challenge in implementing simulated annealing in this
context is defining a proposal function for landscape modifications. At
the moment this is done by computing the minimum-spanning-tree (MST) of
the spatial nodes (locations where occurrence has been observed), and
then proposing corridors that only connect nodes that are already
connected in the MST. The primary reason for doing this is the cut down
the size of the solution space to enable quicker convergence, although
the final software that implements this algorithm will enable
alternative methods of proposing modifications.

The goal output of this chapter is not only to provide a set of discrete
corridor options, but also to rank the cells in the raster by priority
based on how many times they are included in the distribution of
``good'' corridors after simulated annealing has converged. Further, the
final component of this chapter is measuring the effect of land-use
change on the robustness of the optimized corridor by simulating various
neutral models of urban and agricultural sprawl, and determining if the
proposed modifications still maximize time-until-extinction when the
landcover in the landscape is not static.

@@

Currently I have an algorithm for proposing landscape modifications and
a simple implementation of simulated annealing. The only gap left is
implementing Circuitscape estimation of resistance surfaces and running
benchmarking tests for the resulting chapter.

@@

The final chapter consists of a collection of modules in the
\texttt{Julia} language for different aspects of community ecology,
including most of the code used for the preceding chapters. Indeed
\texttt{MetacommunityDynamics.jl} (MCD.jl) is at the center of this set
of tools, but due to the nature of the Julia language, MCD.jl is
inter-operable with several existing packages within the
\texttt{EcoJulia} organization, including several to which I have
contributed. We need a software library like this to generate synthetic
data from a \emph{known} set of mechanisms and parameters to test our
methods for parameter inference and forecasting on this \emph{known}
system to assess the effectiveness of these inference and forecasting
methods.

@@

In fig.~\ref{fig:foodwebtraj} we see a sample output of simulated
food-web dynamics for a metaweb of 100 species generated using the
minimum-potential-niche model (Allesina \emph{et al.} 2008) with
connectance \(C=0.05\) and forbidden-link probability of \(0.5\). The
dynamics change according to a Lotka-Volterra functional response,
dispersal with distance inversely proportional to trophic-level, linear
mortality, and logistic growth for any species at the producer
trophic-level.

@@

The software as it exists is capable of simulating the biomass dynamics
of arbitrarily large food-webs using Lotka-Volterra, Holling Type-II, or
Holling Type-III functional responses. It currently has methods to
implement Gaussian drift, and various forms of dispersal via
\texttt{Dispersal.jl}. Also functional are occupancy dynamics for Levins
metapopulations (Levins 1969), and spatially explicit Hanski-Ovaskainen
metapopulations (Hanski \& Ovaskainen 2000; Ovaskainen \emph{et al.}
2002). This is most of what needs to exist for the preceding chapters.
In-progress functionality includes selection (which affects growth-rate)
on arbitrary environmental variables in progress, as well as traits.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Allesina2008GenMod}{}%
Allesina, S., Alonso, D. \& Pascual, M. (2008). A General Model for Food
Web Structure. \emph{Science}, 320, 658--661.

\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
Banville, F., Vissault, S. \& Poisot, T. (2021). Mangal.jl and
EcologicalNetworks.jl: Two complementary packages for analyzing
ecological networks in Julia. \emph{Journal of Open Source Software}, 6,
2721.
